\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2019-08-21 gwnnDraft02.tex
light drafting in sections 2-4

We consider the following Lagrangian as in~\cite{Arkani-Hamed:2016rle}:
\begin{equation}
{\cal L} = \sum_{i=-N/2}^{N/2} {\cal L}_i,
\end{equation}
with ${\cal L}_0 = {\cal L}_{\rm SM}$ being the Standard Model Lagrangian, and ${\cal L}_i$ being a copy of the SM Lagrangian with different fields, but with all dimensionless parameters the same. Each of the Lagrangians does contain a dimension operator:
\begin{equation}
{\cal L}_i \subset  - \left(m_H^2\right)_i H_i^\dagger H_i  
\end{equation}
where $H_i$ is a Higgs field in each sector, and the mass term is parametrically given by
\begin{equation}\label{eqn:massParam}
\left(m_H^2\right)_i \sim -\frac{\Lambda_H^2}{N}(2i+r),
\end{equation}
where $\Lambda$ is some high scale cutoff, $N$ is the number of sectors, and $r$ mass parameter in the SM in units of $\Lambda_H^2/N$. This setup solves the hierarchy problem if $r \sim \mathcal{O}(1)$\footnote{Constraints require $r$ to be somewhat smaller than 1.} and we the Higgs masses in the different sectors as randomly distributed~\cite{Arkani-Hamed:2016rle}. We have taken for simplicity that there equal numbers sectors with positive and negative $m_H^2$, but this is not nessesary. This Nnaturalness framework can be generalized: the various sectors can possess a wide range of particle content that can be freely selected by the model builder. The one exception to this is that ``our" sector must consist of the Standard Model.

From the above Lagrangians, the Higgs in sectors with $i \geq 0$ will get a vev given by:
\begin{equation}\label{eqn:vevs}
v^i = m^i_H/\sqrt{\lambda_i} ~ \sqrt{\frac{2i + r}{N}}\left(\frac{\Lambda_H}{\lambda}\right),
\end{equation}
$\lambda_i$ is the quartic coefficient of the scalar potential and is the same across all sectors. 
%The parameter $r$ is a way to adjust the tuning of the model: adjusting $r$ away from 1 (which leads to purely uniform spacing) allows us to make a larger gap between our sector and the next one. Within this work, we looked at variants on this sort of tuning including non-uniform distributions and clustered sector effects; these ideas are discussed in Sec. \ref{sec:gw}.   
%
%
%
%
%$N$naturalness is a model framework that looks to solve the hierarchy problem through the introduction of $N$ mutually non-interacting sectors \cite{Arkani-Hamed:2016rle}. The framework itself is very general: the various sectors can possess a wide range of particle content that can be freely selected by the model builder. The one exception to this is that ``our" sector must consist of the Standard Model. For simplicity, the original paper takes all additional new sectors to be copies of the SM (with the same gauge groups and Yukawa structure); however the Higgs mass parameters of each sector are allowed to taken on values distributed between $-\Lambda_H^2$ and $\Lambda^2_H$, with $\Lambda_H$ being the scale that cuts off quadratic divergences.
%
%This construction leads to sectors that are automatically accidentally tuned to the $1/N$ level, leading to a sector with a mass parameter $m_H^2 \sim \Lambda_H^2/N$. The sector with the smallest non-zero vacuum expectation value (vev) as ``our" sector --- the SM sector. It should be noted that the values of $m_H^2$ pass through zero, resulting in effectively 2 types of sectors: ``standard" sectors like our own with a negative Higgs mass parameter squared, possess a vev, and exhibit electroweak symmetry breaking and ``exotic" sectors with a positive Higgs mass parameter squared that thus feature no electroweak symmetry breaking and a zero vev. 
%
%Keeping with the original convention, we write our mass parameter as
%\begin{equation}\label{eqn:massParam}
%\left(m_H^2\right)_i = -\frac{\Lambda_H^2}{N}(2i+r),
%\end{equation}
%where $i$ denotes the sector ($-\frac{N}{2}\leq i \leq \frac{N}{2}$). $i = 0$ is the sector that possess the smallest vev and is thus ``our" SM sector. \textcolor{green}{Write an equation that shows how the vev scales with $i$.} The vev for each standard sector is given explicitly by
%\begin{equation}\label{eqn:vevs}
%v^i = m^i_H/\sqrt{\lambda_i} = \sqrt{\frac{2i + r}{N}}\left(\frac{\Lambda_H}{\lambda}\right),
%\end{equation}
%$\lambda_i$ is the quartic coefficient of the scalar potential and is the same across all sectors. The parameter $r$ is a way to adjust the tuning of the model: adjusting $r$ away from 1 (which leads to purely uniform spacing) allows us to make a larger gap between our sector and the next one. Within this work, we looked at variants on this sort of tuning including non-uniform distributions and clustered sector effects; these ideas are discussed in Sec. \ref{sec:gw}.   
%
%As mentioned above, half of all sectors in this model feature a Higgs mass parameter such that $m_H^2 < 0$. 
These ``standard sectors'' feature electroweak symmetry breaking just like in the SM, however the vevs produced scale with the changing mass parameter: $v_i \sim v_{\rm SM}\sqrt{i}$. As the masses for particles in these sectors scale with the vev, this leads to the masses of the $W$, $Z$ and fermions within each sector increasing with $i$. The consequences of this on the confinement scale of QCD in the $i \geq 1$ sectors is further discussed in Sec. \ref{sec:dQCD}.

The remaining ``exotic sectors'' within this model provide a radical departure from our own. $m_H^2 > 0$ leads to no vev for the Higgs, and electroweak symmetry is only broken at very low scales due to the phase transition from free quarks to confinement at the QCD scale $\Lambda_{QCD}$~\cite{}, and the masses of the $W$ and $Z$ are comparable to those of QCD resonances. The mass of fundamental fermions are produced via four-fermion interactions generated after integrating out the SU$(2)$ Higgs multiplet. This leads to very light fermions: 
\begin{equation}\label{eqn:exoticFM}
m_f \sim y_f y_t \Lambda_{QCD}^3/(m^2_H)_i \leq 100 \,\ev,
\end{equation}
with $y_f$ representing the Yukawa coupling to fermion $f$. As we will see, the extremely light quarks that appear in these sectors dramatically change the nature of the QCD phase transition --- unlike the SM, the transition is strongly first order. Again, this is further developed in Sec. \ref{sec:dQCD}. Crucially, this results in the production of gravitational waves. This is the physical signature we're interested in exploring within this paper; the calculation and results are presented in Sec. \ref{sec:gw}.

\subsection{Reheating N Sectors}

\begin{figure*}[tb]
\centering
\begin{minipage}[c]{\textwidth}
\includegraphics[width=0.4\textwidth]{standardDecay.eps}
\includegraphics[width=0.4\textwidth]{exoticDecay.eps}
\end{minipage}
\hfill
\caption{Feynman diagrams for primary reheaton decays assuming $m_{\phi} \ll m_H$. The left decay is the primary decay for standard sectors ($\langle H\rangle\neq 0$), with $f$ representing all possible fermions. The right decay occurs in exotic sectors ($\langle H\rangle = 0$).
}
\label{fig:reheatonDecays}
\end{figure*}

A key issue within $N$naturalness is how to predominantly gift energy density to our own sector so as to not be immediately excluded by number of effective neutrinos ($N_{eff}$) bounds. This is done through the introduction of a post-inflationary field called the ``reheaton". After inflation, the reheaton field possess the majority of the energy density of the Universe. Although this field can generically be either by bosonic or fermionic, we reduce our scope to a scalar reheaton $\phi$. Our focus is primarily the production of gravitational waves from multiple sectors: \textcolor{yellow}{What does the following mean?} a fermion reheaton doesn't change the scaling of the energy density of the exotic sectors and thus doesn't affect expected gravitational wave profiles.

Ultimately, after examining the gravitational wave case produced by standard $N$naturalness, we generalize our results to include the (rather unnatural) case where sectors are reheated ``randomly" as opposed to in relation to the higgs mass parameter. This allows us to explore more general models with multiple dark sectors at a huge range of scales. For these models, the reheating mechanism remains undefined and the results of this section are completely irrelevant.

The appropriate Lagrangian for a scalar reheaton $\phi$ is: 
\begin{equation}\label{eqn:nnLagrangian}
\cL_\phi \supset -a\phi\sum_i\vert H_i\vert^2 - \frac{1}{2}m^2_\phi \phi^2.
\end{equation}
Note that the cross-quartic couplings are absent, suppressed by a very small coupling. Effective Lagrangians for the two different types of sectors present in this theory can be obtained through the integrating out of the Higgs bosons in every sector:
\begin{equation}\label{eqn:nnEffLag}
\begin{split}
\cL_\phi^{v \neq 0} &\supset C_1 a y_q\frac{v}{m_h^2}\phi q q^c,
\\
\cL_\phi^{v = 0} &\supset C_2 a \frac{g^2}{16\pi^2 m_H^2}\phi W_{\mu\nu}W^{\mu\nu},
\end{split}
\end{equation}
\textcolor{green}{Should the first equation go like $1/m_h$ (not squared)? It does --- v scales identically to $m_h$}with $C_i$ representing numerical coefficients, $g$ the weak coupling constant, and $W^{\mu\nu}$ the SU$(2)$ field strength tensor.

Immediately from Eq. (\ref{eqn:nnEffLag}), we can see that the matrix element for decays into standard sectors is inversely proportional to that sectors higgs mass, $\cM_{m_H^2 < 0} \sim 1/m_{h_i}$. The Feynman diagram for this process is presented in Fig. \ref{fig:reheatonDecays}. The loop decay of $\phi \rightarrow \gamma\gamma$ is always sub-leading and can be neglected. It should be noted that as one goes to sectors with larger and larger vevs, the increasing mass of the fermions ($m_f \sim v_i \sim v_{SM}\sqrt{i}$) eventually leads to situations where the decay to two on-shell bottom or charm quarks is kinematically forbidden, $m_\phi < 2 m_q$. \textcolor{green}{Put an equation sketching how $m_f$ scales with $m_h$} For sectors where this kinematic threshold is passed for charm quarks, contributions to cosmological observables can be safely ignored. All in all, we end up with a decay width that scales as $\Gamma_{m_H^2<0} \sim 1/m_h^2$. Since we can expect energy density to be proportional to the decay width, $\frac{\rho_i}{\rho_{SM}} \approx \frac{\Gamma_i}{\Gamma_{SM}}$, this indicates that energy density of standard sectors falls:
\begin{equation}\label{eqn:edSS}
\rho_i \sim r_{s}\frac{\rho_{SM}}{i}
\end{equation} 
with $r_s$ being the ratio of the energy density of the first additional standard sector over the energy density of our sector.

For the exotic sectors, Eq. (\ref{eqn:nnEffLag}) indicates a matrix element scaling $\cM_{m_H^2>0} \sim 1/m_{H_i}^2$. As the Feynman diagram (Fig.~\ref{fig:reheatonDecays}) shows, the reheaton decays to exotic sectors are loop suppressed, leading to a significantly lower energy density than the standard sectors. Both the decay width and energy density for these sectors scale as: $\Gamma_{m_H^2>0} \sim \rho_i \sim 1/m_H^4 \sim 1/i^2$.

As a final note, the reheating temperature of our sector, $T_{RH}$, \textcolor{yellow}{What is this parameter exactly? The reheat temperature of the SM? Also, is this effect discussed in the original NNaturalness paper? Yes to both.} has an upper bound on the order of the weak scale. If this bound is not observed, the SM Higgs mass would have major thermal corrections --- leading to the branching ratios into other sectors being problematically large \citep{Arkani-Hamed:2016rle}. In accordance with this restriction, this project only examines reheating temperatures at or below $100$ GeV.

\section{Dark QCD}
\label{sec:dQCD}

\begin{figure*}[tb]
\centering
\begin{minipage}[c]{\textwidth}
\includegraphics[width=1.0\textwidth]{columbia.png}
\end{minipage}
\hfill
\caption{Two possible scenarios for the Columbia Plot with the up and down quark mass taken to be identical. The ``physical point" is the Standard Model. Plot from \cite{Cuteri:2017zcb}
}
\label{fig:columbia}
\end{figure*}

Due to the confining nature of QCD, the exact nature of the phase transition is often difficult to ascertain analytically and requires the study of lattice simulations. The results of these studies on the nature of QCD phase transitions is summarized in the so-called ``Columbia" plot (Fig. \ref{fig:columbia}). As seen in the above figure (and explicitly demonstrated through lattice studies \cite{lattice}), our sector is solidly in the weak-cross over area. In order to observe the strong first order transition required for gravitational wave production, at least some of the additional sectors present must feature quarks in either the heavy quark (pure Yang-Mills) or massless limits.

In the Yang-Mills limit we see $m_{u,d,s} \rightarrow \infty$. This leads to the presence of a global $Z_3$ centre symmetry that is broken at high temperatures but restored low temperatures. Ultimately, this restoration results in a first order phase transition \cite{SVETITSKY1982423}. On the other end of the spectrum, ``massless" quark theories ($m_q \ll \Lambda_{QCD}$ for all quarks) also contain strong phase transitions due to the breakdown of SU$(N)\,\times$ SU$(N)$ chiral symmetry \cite{Pisarski:1983ms}. Both the Yang-Mills and massless limit arguments can be generalized to SU$(N \geq 3)$: coupling both with the additional restriction that we avoid the conformal window indicates that strong first order phase transitions can occur if we have $n = 0$ (pure Yang-Mills) or $3 \leq n < 4N$ (light quark limit) very light flavours.

For $N$naturalness (or other mirror sector models), standard sectors with higher vevs possess more massive quarks; eventually, sectors with large enough vevs push us into the upper-right corner. Conversely, exotic sectors with zero vev feature incredibly light quarks and drop us into the bottom-left corner. In order to figure out the specifics of each sector, we follow  the same procedure as \cite{Cui:2011wk} but generalize it to an arbitrary number of additional sectors. First, due to the parameters of each sector being taken to be identical save for the higgs mass squared (thus $v \neq v_i$, where $v$ is the SM vev), we assume that the strong coupling of every sector is identical at some high scale. 
Explicitly, above the top quark mass scale $\Lambda_{QCD} = 89 \pm 5$ in $\overline{MS}$ \cite{PhysRevD.98.030001}. Since the running of the gauge coupling at scale $\mu$ is given by 
\begin{equation}\label{eqn:QCDrunning}
\alpha_s (\mu) = \frac{2\pi}{11-\frac{2n_f}{3}}\frac{1}{\ln{\mu/\Lambda}}
\end{equation} 
where $n_f$ is the number of quark flavours and $\Lambda$ is the confinement scale $\Lambda_{QCD}$ \cite{Peskin:1995ev}.

For any other sector $i$ we write out the expression as: 
\begin{equation}\label{eqn:QCDrunningi}
\alpha_{s}^i (\mu) = \frac{2\pi}{11-\frac{2n^i_f}{3}}\frac{1}{\ln{\mu/\Lambda^i}},
\end{equation}
featuring the equivalent definitions for the additional sectors. Because we've set the strong couplings equal at high scales, $\Lambda = \Lambda^i$ for all $i$ at scales above the top quark mass for the heaviest sector. However, since the masses of the quarks in each sector are different, we end up with a unique running of the coupling for each sector. At every quark mass threshold for a given sector, we match the coupling strengths above and below the threshold and determine the new $\Lambda^i$ for the lower scale:
\begin{equation}
\alpha_s^{i(5)} = \alpha_s^{i(6)}
\end{equation} 
and thus
\begin{equation}
\Lambda_{(5)}^i = (m_t^{i})^{2/23}(\Lambda_{(6)}^i)^{21/23}.
\end{equation}

Suppressing the $i$s for cleanliness, we can arrive at similar relations at the bottom and charm thresholds
\begin{equation}
\begin{split}
\Lambda_{(4)} = (m_b)^{2/25}(\Lambda_{(5)})^{23/25},
\\
\Lambda_{(3)} = (m_c)^{2/23}(\Lambda_{(4)})^{25/27}.
\end{split}
\end{equation}
These can be combined to show that
\begin{equation}
\Lambda_{(3)} = (m_t m_b m_c)^{2/27}(\Lambda_{(6)})^{21/27}.
\end{equation}
This type of matching procedure can be done as many times as necessary for a given sector. The process terminates when $\Lambda_i$ for a given scale is larger than the next quark mass threshold (i.e running the scale down arrives at the $\Lambda_{QCD}$ phase transition before reaching the next quark mass scale).

@@

As mentioned in the previous sections (see Eq.~(\ref{eqn:vevs})),\textcolor{green}{refer to a particular equation} for standard sectors with increasing index $i$ the vevs of said sectors increase $v_i\propto \sqrt{i}$. This leads to increasingly heavy particle spectra for higher sectors --- eventually leading to sectors that are essentially pure Yang-Mills that featuring strong first order phase transitions. This, of course, begs the question: at what index $i$ do said phase transitions begin?

The actual boundary of region of interest on the Columbia plot is not precisely known and further lattice simulations are required to determine exactly where in parameter space cross-over transitions end and first order transitions begin. In lieu of such studies, we parameterize our ignorance in the form of a ``slop factor" $\kappa$. Assuming the up and down quarks of each sector to be of roughly the same scale, we determine what sector $i$ features $\kappa m_u^i \approx \Lambda_{QCD}$.

Using the methods outlined in the prior section we determine $\Lambda_{QCD}$ to have a relevant value of 
\begin{equation}\label{eqn:lambda2}
\Lambda^i_{(2)} = (m_s^i m_c^i m_b^i m_t^i)^{2/29}(\Lambda^i_{(6)})^{21/29}
\end{equation} 
at the energy scale we're interested in. $\Lambda^i_{(6)}$ is identical for all sectors and is taken to have a standard model value of $\Lambda^{(6)}_{MS} =(89 \pm 6)\, \mev$\cite{PhysRevD.98.030001}\textcolor{green}{what is that value?}. Rewriting Eq.~(\ref{eqn:lambda2}) in terms of standard model variables, 
\begin{equation}\label{eqn:lambda2adj}
\Lambda^i_{(2)} = (m_s m_c m_b m_t i^2)^{2/29}(\Lambda_{(6)})^{21/29}.
\end{equation}
Setting this equal to some multiple of the quark mass of sector $i$,
\begin{equation}
\kappa m^i_u = \kappa m_u \sqrt{i} = (m_s m_c m_b m_t i^2)^{2/29}(\Lambda_{(6)})^{21/29}.
\end{equation}
This can be solved for $i$:
\begin{equation}\label{eqn:critIndex}
i = \frac{(m_s m_c m_b m_t)^{4/25}(\Lambda_{(6)})^{42/25}}{(\kappa m_u)^{58/25}}.
\end{equation}

The critical index demonstrates that even in the most extreme cases, where first order phase transitions occur with the up quark an order of magnitude less than $\Lambda_{QCD}$, the critical index is still $\sim 1000$. Looking back to the energy density scaling in Eq.~(\ref{eqn:edSS})\textcolor{green}{refer to a particular equation} we see that, within the context of $N$naturalness, the energy density of these sectors is at most $10^{-4}$ \textcolor{green}{why not $10^{-3}$?} that of our sector. As a result, \textcolor{green}{As we will show below?} the gravitational waves generated by the standard large $i$ sectors in $N$naturalness will not generate detectable signatures.

However, if we move away from the standard $N$naturalness reheating mechanism and begin exploring mirror sectors with large vevs with $\cO( >10\%)$ of our energy density $\rho_{SM}$, we can have sectors with relatively high dark QCD scales that produce detectable gravitational waves. These sectors feature no baryons and, as such, produce runaway bubbles similar to the exotic sectors in standard $N$naturalness. From Eq.~(\ref{eqn:lambda2}) we can determine the confinement scale of an arbitrary mirror sector --- the maximum possible given a $10^{16}$ limit \cite{Arkani-Hamed:2016rle} \textcolor{yellow}{Is this actually an upper value of $m_H^2$ at the gut scale? The more sectors added in, the lower the scale of strong gravitational coupling. With more sectors than $10^16$, the top sectors start having vevs higher than the gravitational scale} of sectors is $\sim 38 \,\gev$. After imposing $N_{eff}$ constraints (as in Sec. \ref{sec:constraints}), the energy density of said sector is limited to $\sim 0.25\rho_{SM}$. The signals of this sector and other test cases like it are explored in Sec. \ref{sec:gw}.

>>>>

We consider the following Lagrangian as in~\cite{Arkani-Hamed:2016rle}:
\begin{equation}
{\cal L} = \sum_{i=-N/2}^{N/2} {\cal L}_i,
\end{equation}
with ${\cal L}_0 = {\cal L}_{\rm SM}$ being the Standard Model Lagrangian, and ${\cal L}_i$ being a copy of the SM Lagrangian with different fields, but with all dimensionless parameters the same. Each of the Lagrangians does contain a dimensionful operator:
\begin{equation}
{\cal L}_i \subset  - \left(m_H^2\right)_i H_i^\dagger H_i  
\end{equation}
where $H_i$ is a Higgs field in each sector, and the mass term is parametrically given by
\begin{equation}\label{eqn:massParam}
\left(m_H^2\right)_i \sim -\frac{\Lambda_H^2}{N}(2i+r),
\end{equation}
where $\Lambda$ is some high scale cutoff, $N$ is the number of sectors, and $r$ mass parameter in the SM in units of $\Lambda_H^2/N$. We view the parameterization of Eq.~(\ref{eqn:massParam}) as a random distribution in theory space up to the cutoff $\Lambda$, therefore this setup solves the hierarchy problem if $r \sim \mathcal{O}(1)$~\cite{Arkani-Hamed:2016rle}\footnote{Constraints require $r$ to be somewhat smaller than 1.} and our sector is the one that that has the smallest absolute value of the Higgs mass parameter. We have taken for simplicity that there equal numbers sectors with positive and negative $m_H^2$, but this assumption does not affect our analysis. This Nnaturalness framework can be generalized: the various sectors can possess a wide range of particle content that can be freely selected by the model builder. The one exception to this is that ``our" sector must consist of the Standard Model.

From the above Lagrangians, the Higgs in sectors with $i \geq 0$ will get a vev given by:
\begin{equation}\label{eqn:vevs}
v^i = \sqrt{-(m_H^2)_i/\lambda_i} \sim \Lambda_H\sqrt{\frac{2i + r}{\lambda N}},
\end{equation}
$\lambda_i$ is the quartic coefficient of the scalar potential and is the same across all sectors, $\lambda_i=\lambda$. This is another way to see how this framework can solve the hierarchy problem: the Higgs vev is parametrically smaller than the cutoff for $N \gg 1$. 
%The parameter $r$ is a way to adjust the tuning of the model: adjusting $r$ away from 1 (which leads to purely uniform spacing) allows us to make a larger gap between our sector and the next one. Within this work, we looked at variants on this sort of tuning including non-uniform distributions and clustered sector effects; these ideas are discussed in Sec. \ref{sec:gw}.   
%
%
%
%
%$N$naturalness is a model framework that looks to solve the hierarchy problem through the introduction of $N$ mutually non-interacting sectors \cite{Arkani-Hamed:2016rle}. The framework itself is very general: the various sectors can possess a wide range of particle content that can be freely selected by the model builder. The one exception to this is that ``our" sector must consist of the Standard Model. For simplicity, the original paper takes all additional new sectors to be copies of the SM (with the same gauge groups and Yukawa structure); however the Higgs mass parameters of each sector are allowed to taken on values distributed between $-\Lambda_H^2$ and $\Lambda^2_H$, with $\Lambda_H$ being the scale that cuts off quadratic divergences.
%
%This construction leads to sectors that are automatically accidentally tuned to the $1/N$ level, leading to a sector with a mass parameter $m_H^2 \sim \Lambda_H^2/N$. The sector with the smallest non-zero vacuum expectation value (vev) as ``our" sector --- the SM sector. It should be noted that the values of $m_H^2$ pass through zero, resulting in effectively 2 types of sectors: ``standard" sectors like our own with a negative Higgs mass parameter squared, possess a vev, and exhibit electroweak symmetry breaking and ``exotic" sectors with a positive Higgs mass parameter squared that thus feature no electroweak symmetry breaking and a zero vev. 
%
%Keeping with the original convention, we write our mass parameter as
%\begin{equation}\label{eqn:massParam}
%\left(m_H^2\right)_i = -\frac{\Lambda_H^2}{N}(2i+r),
%\end{equation}
%where $i$ denotes the sector ($-\frac{N}{2}\leq i \leq \frac{N}{2}$). $i = 0$ is the sector that possess the smallest vev and is thus ``our" SM sector. \textcolor{green}{Write an equation that shows how the vev scales with $i$.} The vev for each standard sector is given explicitly by
%\begin{equation}\label{eqn:vevs}
%v^i = m^i_H/\sqrt{\lambda_i} = \sqrt{\frac{2i + r}{N}}\left(\frac{\Lambda_H}{\lambda}\right),
%\end{equation}
%$\lambda_i$ is the quartic coefficient of the scalar potential and is the same across all sectors. The parameter $r$ is a way to adjust the tuning of the model: adjusting $r$ away from 1 (which leads to purely uniform spacing) allows us to make a larger gap between our sector and the next one. Within this work, we looked at variants on this sort of tuning including non-uniform distributions and clustered sector effects; these ideas are discussed in Sec. \ref{sec:gw}.   
%
%As mentioned above, half of all sectors in this model feature a Higgs mass parameter such that $m_H^2 < 0$. 
The ``standard sectors'' with $i>0$ feature electroweak symmetry breaking just like in the SM, however the vevs produced scale with the changing mass parameter: $v_i \sim v_{\rm SM}\sqrt{i}$. As the masses scale with the vev, this leads to the masses of the $W$, $Z$ and fermions within each sector increasing with $i$. The consequences of this on the confinement scale of QCD in the $i \geq 1$ sectors is further discussed in Sec. \ref{sec:dQCD}.

The ``exotic sectors'' with $i<0$ provide a radical departure from our own. $m_H^2 > 0$ leads to no vev for the Higgs, and electroweak symmetry is only broken at very low scales due to the phase transition from free quarks to confinement at the QCD scale $\Lambda_{QCD}$~\cite{}, and the masses of the $W$ and $Z$ are comparable to those of QCD resonances. The mass of fundamental fermions are produced via four-fermion interactions generated after integrating out the SU$(2)$ Higgs multiplet. This leads to very light fermions: 
\begin{equation}\label{eqn:exoticFM}
m_f \sim y_f y_t \Lambda_{QCD}^3/(m^2_H)_i \leq 100 \,\ev,
\end{equation}
with $y_f$ representing the Yukawa coupling to fermion $f$. As we will see, the extremely light quarks that appear in these sectors dramatically change the nature of the QCD phase transition --- unlike the SM, the transition is strongly first order. Again, this is further developed in Sec. \ref{sec:dQCD}. Crucially, this results in the production of gravitational waves. This is the physical signature we're interested in exploring within this paper; the calculation and results are presented in Sec. \ref{sec:gw}.

\section{QCD Phase Transition}
\label{sec:dQCD}
%
%\begin{figure*}[tb]
%\centering
%\begin{minipage}[c]{\textwidth}
%\includegraphics[width=1.0\textwidth]{columbia.png}
%\end{minipage}
%\hfill
%\caption{Two possible scenarios for the Columbia Plot with the up and down quark mass taken to be identical. The ``physical point" is the Standard Model. Plot from \cite{Cuteri:2017zcb}
%}
%\label{fig:columbia}
%\end{figure*}

We now study the nature of the QCD phase transition across the different sectors. Due to the confining nature of QCD, the exact nature of the phase transition is often difficult to ascertain analytically and requires the study of lattice simulations. In the SM, it is known that the phase transition is a crossover and does lead to gravitational wave signals~\cite{}. In the general case \draftnote{Is this specific to three colours?}, the phase transition can be strongly first order in two regimes~\cite{lattice,SVETITSKY1982423,Pisarski:1983ms}:
\begin{itemize}
\item Three or more light flavours.
\item No light flavours. 
\end{itemize}
Light means mass small compared to the confinement scale $\Lambda_{QCD}$, but what that means quantitatively is not precisely determined. In the SM, the up and down quarks are light, but the strange is not sufficiently light for an SFOPT.

% The results of these studies on the nature of QCD phase transitions is summarized in the so-called ``Columbia" plot (Fig. \ref{fig:columbia}). As seen in the above figure (and explicitly demonstrated through lattice studies \cite{lattice}), our sector is solidly in the weak-cross over area. In order to observe the strong first order transition required for gravitational wave production, at least some of the additional sectors present must feature quarks in either the heavy quark (pure Yang-Mills) or massless limits.    
%
%In the Yang-Mills limit we see $m_{u,d,s} \rightarrow \infty$. This leads to the presence of a global $Z_3$ centre symmetry that is broken at high temperatures but restored low temperatures. Ultimately, this restoration results in a first order phase transition \cite{SVETITSKY1982423}. On the other end of the spectrum, ``massless" quark theories ($m_q \ll \Lambda_{QCD}$ for all quarks) also contain strong phase transitions due to the breakdown of SU$(N)\,\times$ SU$(N)$ chiral symmetry \cite{Pisarski:1983ms}. Both the Yang-Mills and massless limit arguments can be generalized to SU$(N \geq 3)$: coupling both with the additional restriction that we avoid the conformal window indicates that strong first order phase transitions can occur if we have $n = 0$ (pure Yang-Mills) or $3 \leq n < 4N$ (light quark limit) very light flavours.

For the standard sectors in our setup, the quark masses increase with increasing vev, so for sufficiently large $i$, all the quarks will be heavier than $\Lambda_{QCD}$,\footnote{$\Lambda_{QCD}$ does vary with $i$, but the sensitivity is very weak as we will see below.} and those large $i$ sectors will undergo an SFOPT if they are reheated above the the confinement scale. 
%For $N$naturalness (or other mirror sector models), standard sectors with higher vevs possess more massive quarks; eventually, sectors with large enough vevs push us into the upper-right corner. 
Conversely, exotic sectors with zero vev feature six very light quarks, so \textit{all} the exotic sectors undergo SFOPT at the temperature of QCD confinement.

We now calculate the QCD confinement scale for each sector following the same procedure as \cite{Cui:2011wk}. First, due to the parameters of each sector being taken to be identical save for the higgs mass squared (thus $v \neq v_i$, where $v$ is the SM vev), we assume that the strong coupling of every sector is identical at some high scale. Using the one-loop running, the $\beta$ function can be solved:
\begin{equation}\label{eqn:QCDrunningi}
\alpha_{s}^i (\mu) = \frac{2\pi}{11-\frac{2n^i_f}{3}}\frac{1}{\ln{\mu/\Lambda^i}},
\end{equation}
where $n_f^i$ is the number of quark flavours with mass less than $\mu/2$ and $\Lambda^i$ is the scale where it would confine if all quarks remain massless. In the SM defined at scales well above all the quark masses, we have $\Lambda_{QCD} = 89 \pm 5$ in $\overline{MS}$~\cite{PhysRevD.98.030001}.

Because we've set the strong couplings equal at high scales, $\Lambda = \Lambda^i$ for all $i$ at high scales for all sectors. However, since the masses of the quarks in each sector are different, we end up with a unique running of the coupling for each sector. At every quark mass threshold for a given sector, we match the coupling strengths above and below the threshold and determine the new $\Lambda^i$ for the lower scale. For example, at the mass of the top quark, we match a five flavour coupling with the six flavour one:
\begin{equation}
\alpha_s^{i(5)}(2 m^i_t) = \alpha_s^{i(6)}(2 m^i_t) 
\end{equation} 
and thus
\begin{equation}
\Lambda_{(5)}^i = (m_t^{i})^{2/23}(\Lambda_{(6)}^i)^{21/23}.
\end{equation}

Suppressing the $i$'s for notational cleanliness, we can arrive at similar relations at the bottom and charm thresholds
\begin{equation}
\begin{split}
\Lambda_{(4)} = (m_b)^{2/25}(\Lambda_{(5)})^{23/25},
\\
\Lambda_{(3)} = (m_c)^{2/23}(\Lambda_{(4)})^{25/27}.
\end{split}
\end{equation}
These can be combined to show that
\begin{equation}
\Lambda_{(3)} = (m_t m_b m_c)^{2/27}(\Lambda_{(6)})^{21/27}.
\end{equation}
This type of matching procedure can be done as many times as necessary for a given sector. The process terminates when $\Lambda_i$ for a given scale is larger than the next quark mass threshold (i.e running the scale down arrives at the $\Lambda_{QCD}$ phase transition before reaching the next quark mass scale).

@@

As shown in Eq.~(\ref{eqn:vevs}), for standard sectors with increasing index $i$ the vevs of said sectors increase $v_i\propto \sqrt{i}$. This leads to increasingly heavy particle spectra for higher sectors --- eventually leading to sectors that are essentially pure Yang-Mills that featuring strong first order phase transitions. This, of course, begs the question: at what index $i$ do said phase transitions begin?

%The actual boundary of region of interest on the Columbia plot is not precisely known and further lattice simulations are required to determine exactly where in parameter space cross-over transitions end and first order transitions begin. In lieu of such studies, we parameterize our ignorance in the form of a ``slop factor" $\kappa$. Assuming the up and down quarks of each sector to be of roughly the same scale, we determine what sector $i$ features $\kappa m_u^i \approx \Lambda_{QCD}$.

Using the methods outlined in the prior section we determine $\Lambda_{QCD}$ to have a relevant value of 
\begin{equation}\label{eqn:lambda2}
\Lambda^i_{(2)} = (m_s^i m_c^i m_b^i m_t^i)^{2/29}(\Lambda^i_{(6)})^{21/29}
\end{equation} 
at the energy scale we're interested in. $\Lambda^i_{(6)}$ is identical for all sectors and is taken to have a standard model value of $\Lambda^{(6)}_{MS} =(89 \pm 6)\, \mev$~\cite{PhysRevD.98.030001}. Rewriting Eq.~(\ref{eqn:lambda2}) in terms of standard model variables, 
\begin{equation}\label{eqn:lambda2adj}
\Lambda^i_{(2)} = (m_s m_c m_b m_t i^2)^{2/29}(\Lambda_{(6)})^{21/29}.
\end{equation}
We take the sector with SFOPT to be the ones when the mass of the up quark, down quark, and QCD phase transition scale are all comparable:
\begin{equation}
m^i_u \sim m_u \sqrt{i} \sim (m_s m_c m_b m_t i^2)^{2/29}(\Lambda_{(6)})^{21/29}.
\end{equation}
This can be solved for $i$:
\begin{equation}\label{eqn:critIndex}
i^c \sim \frac{(m_s m_c m_b m_t)^{4/25}(\Lambda_{(6)})^{42/25}}{( m_u)^{58/25}} \sim 10^5
\end{equation}
%
As we will see in Sec.~\ref{sec:reheat}, In the original Nnaturalness setup~\cite{Arkani-Hamed:2016rle}, the energy dumped into the $i$th sector scales as $i^{-8}$ \draftnote{check}, so there will will not be enough energy in the sectors with $i>I^c$ to see a signature of these phase transitions.

%This critical index demonstrates that even in the most extreme cases, where first order phase transitions occur with the up quark an order of magnitude less than $\Lambda_{QCD}$, the critical index is still $\sim 1000$. Looking back to the energy density scaling in Eq.~(\ref{eqn:edSS})\textcolor{green}{refer to a particular equation} we see that, within the context of $N$naturalness, the energy density of these sectors is at most $10^{-4}$ \textcolor{green}{why not $10^{-3}$?} that of our sector. As a result, \textcolor{green}{As we will show below?} the gravitational waves generated by the standard large $i$ sectors in $N$naturalness will not generate detectable signatures.

However, if we move away from the original $N$naturalness reheating mechanism and begin exploring mirror sectors with large vevs and with relative energy densities $\rho_i / \rho_{SM} \sim 10\%$, a possibility allowed by current constraints, we can have sectors with relatively high dark QCD scales that produce detectable gravitational waves. From Eq.~(\ref{eqn:lambda2}) we can determine the confinement scale of an arbitrary mirror sector. If we take Higgs vevs as high as the GUT scale $\sim 10^{16}$ GeV, then we can gave confinement scales as high as $\sim 38 \,\gev$.  The signals of this sector and other test cases like it are explored in Sec. \ref{sec:gw}.

\draftnote{Move the below two statements to the contraints and GW sections respectively}

After imposing $N_{eff}$ constraints (as in Sec. \ref{sec:constraints}), the energy density of said sector is limited to $\sim 0.25\rho_{SM}$.

These sectors feature no baryons and, as such, produce runaway bubbles similar to the exotic sectors in standard $N$naturalness.
